---
title: "Assignment 2"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(MASS)
library(dplyr)
library(lubridate)
library(surveillance)
library(ggplot2)
library(shiny)
library(ciTools)
source('./R/glrpoisApp.R')
source('./R/surv.ds.convert.R')

#source('./R/Prep_raw_data.R')

```


## Import Data

This dataset (a1) contains the number of people visiting Emergency Departments in Manhattan, NY for diarrhea.

```{r}
a2 <- readRDS('./Data/65p_diarrhea_manhattan.rds') %>%
  filter(date<'2020-03-01') %>%
  mutate(day_of_week = as.factor(weekdays(date))) #day of week

```


## Plot the time series for Children under 5
```{r}

```

## Harmonic regression

Our goal is to establish a prospective aberration detective algorithm for diarrhea in kids. Let's first focus on the period before the pandemic, which is subset and saved in data frame a2 below:

Create the harmonic variables you need for a harmonic regression. Remember, this is daily data


Make sure to sort by date, then create an index, and harmonic terms
```{r}

a3 <- a2 %>%
  

```


Fit your Poisson regression model and negative binomial--which will you choose?

```{r}

mod1 <- glm(XX~YY, data=a3, family='poisson')

summary(mod1)


mod2 <- glm.nb(XX~YY, data=a3)

summary(mod2)
```

Plot your output observed and fitted values

```{r}
pred_mod2 <-  add_pi(a3, fit=mod1, yhatName = "pred", alpha=0.05 )

p1 <- ggplot(a3, aes(x=date, y=count)) +
  geom_line() +
  theme_classic()+
  geom_ribbon(data=pred_mod2, 
          aes(ymin=LPB0.025,ymax=UPB0.975), fill="blue", alpha=0.5) +
  geom_line(data=pred_mod2, aes(x=date, y=pred), col='gray', lty=2)

p1
```
## Nowcasting

Line list data, Two columns: one for case detection date, one for date of reporting

```{r}
ds1 <-  readRDS('Data/reported_cases.rds')

```


Calculate the delay distribution (what proportion of cases were reported after day 1,2,3,4...), and plot the delay distribution and the cumlative delays (what proportion had been reported by day 1, day 2, day 3)
```{r}


```

Create a time series based on date of diagnosis
```{r}

```

Create a time series based on date of report
```{r}

```

Plot these two time series
```{r}

```


## Correct the observed data based on date of diagnosis using the delay distribution

```{r}

```

#Plot the reported and corrected cases

```{r}

```
compare to actual complete cases (not yet fully observed)

```{r}
#The actual full dataset that is observed retrospectively
 full_ds <- readRDS( 'Data/reported_cases_full_final.rds') %>%
  group_by(date_of_diagnosis) %>%
  summarize(N_full=n())



```

## Reporting triangle approach

## Generate a reporting triangle

```{r}
# Reporting triangle
# Reporting triangle on BY March 19
reporting.triangle.mar.19 <- ds1 %>% 
  mutate( diff = interval(date_of_diagnosis,date_of_report  ),
          delay = diff %/% days(1),
          date_of_diagnosis=as.factor(date_of_diagnosis)) %>%
  select(date_of_diagnosis, delay) %>%
  group_by(date_of_diagnosis,delay) %>%
  summarize(N=n()) %>%
  ungroup() %>%
  tidyr::complete(date_of_diagnosis, delay, fill=list(N=0)) %>%
  ungroup() %>%
  mutate(date_of_diagnosis=as.Date(date_of_diagnosis)) %>%
  filter(date_of_diagnosis + delay <='2022-01-18' )%>%
  arrange(date_of_diagnosis, delay) %>%
  reshape2::dcast(date_of_diagnosis~delay, value.var='N')


  write.csv(reporting.triangle.jan.18, 'Data/reporting.triangle_jan18.csv')
```





