---
title: "Assignment 2"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(MASS)
library(dplyr)
library(lubridate)
library(surveillance)
library(ggplot2)
library(shiny)
library(ciTools)
source('./R/glrpoisApp.R')
source('./R/surv.ds.convert.R')

#source('./R/Prep_raw_data.R')

```


## Import Data

This dataset (a1) contains the number of people visiting Emergency Departments in Manhattan, NY for diarrhea.

```{r}
a2 <- readRDS('./Data/65p_diarrhea_manhattan.rds') %>%
  filter(date<'2020-03-01') %>%
  mutate(day_of_week = as.factor(weekdays(date))) #day of week

```


## Plot the time series for Children under 5
```{r}
 ggplot(a2, aes(x=date, y=count)) +
  theme_classic() +
  geom_line()
```

## Harmonic regression

Our goal is to establish a prospective aberration detective algorithm for diarrhea in kids. Let's first focus on the period before the pandemic, which is subset and saved in data frame a2 below:

Create the harmonic variables you need for a harmonic regression. Remember, this is daily data


Make sure to sort by date, then create an index, and harmonic terms
```{r}

a3 <- a2 %>%
  arrange(date) %>%
  mutate( t=row_number(),
          sin1= sin(2*pi*t/365),
          cos1= cos(2*pi*t/365),

          sin2= sin(2*pi*t*2/365),
          cos2= cos(2*pi*t*2/365),

          sin3= sin(2*pi*t*3/365),
          cos3= cos(2*pi*t*3/365),
          
          
          )


```


Fit your Poisson regression model and negative binomial--which will you choose?

```{r}

mod1 <- glm(count~sin1 +cos1+ sin2 +cos2 +sin3+cos3+t +day_of_week, data=a3, family='poisson')

summary(mod1)


mod2 <- glm.nb(count~sin1 +cos1+ sin2 +cos2 +sin3+cos3+t + day_of_week, data=a3)

summary(mod2)
```

Plot your output observed and fitted values

```{r}
pred_mod2 <-  add_pi(a3, fit=mod1, yhatName = "pred", alpha=0.05 )

p1 <- ggplot(a3, aes(x=date, y=count)) +
  geom_line() +
  theme_classic()+
  geom_ribbon(data=pred_mod2, 
          aes(ymin=LPB0.025,ymax=UPB0.975), fill="blue", alpha=0.5) +
  geom_line(data=pred_mod2, aes(x=date, y=pred), col='gray', lty=2)

p1
```
## Nowcasting

Line list data, Two columns: one for case detection date, one for date of reporting

```{r}
ds1 <-  readRDS('Data/reported_cases.rds')

```


Plot the delay distribution (what proportion of cases were reported after day 1,2,3,4...)
```{r}
proportion_reporting <- ds1  %>%  
  mutate(delay = as.numeric(date_of_report -date_of_diagnosis )) %>% 
  group_by(delay) %>%
  summarize(N=n()) %>%
  ungroup() %>%
  tidyr::complete(delay=1:30) %>%
  mutate(prop= N/sum(N, na.rm=T),
         prop = if_else(is.na(prop),0,prop) ,
         cum_prop=cumsum(prop))

ggplot(test1, aes(x=delay, y=prop)) +
  geom_line()+
  theme_classic()+
    ggtitle('Proportion Reported by delay day')


ggplot(test1, aes(x=delay, y=cum_prop)) +
  geom_line()+
  theme_classic()+
  ggtitle('Cumulative Proportion Reported by delay day')

```

Create a time series based on date of diagnosis
```{r}
ts_date_dx <- ds1 %>% 
  group_by(date_of_diagnosis) %>%
  summarize(N_date_dx = n() ) %>%
  rename(date=date_of_diagnosis)
```

Create a time series based on date of report
```{r}
ts_date_report <- ds1 %>% 
  group_by(date_of_report) %>%
  summarize(N_date_report = n() ) %>%
  rename(date=date_of_report)
```

```{r}
ggplot(ts_date_dx) +
  geom_line(aes(x=date, y=N_date_dx)) +
  geom_line(data=ts_date_report,aes(x=date, y=N_date_report),color='red') +
 theme_classic() 
  

```


## Generate a reporting triangle

```{r}
# Reporting triangle
# Reporting triangle on BY March 19
reporting.triangle.mar.19 <- ds1 %>% 
  mutate( diff = interval(date_of_diagnosis,date_of_report  ),
          delay = diff %/% days(1),
          date_of_diagnosis=as.factor(date_of_diagnosis)) %>%
  select(date_of_diagnosis, delay) %>%
  group_by(date_of_diagnosis,delay) %>%
  summarize(N=n()) %>%
  ungroup() %>%
  tidyr::complete(date_of_diagnosis, delay, fill=list(N=0)) %>%
  ungroup() %>%
  mutate(date_of_diagnosis=as.Date(date_of_diagnosis)) %>%
  filter(date_of_diagnosis + delay <='2022-01-18' )%>%
  arrange(date_of_diagnosis, delay) %>%
  reshape2::dcast(date_of_diagnosis~delay, value.var='N')


  write.csv(reporting.triangle.jan.18, 'Data/reporting.triangle_jan18.csv')
```
## Correct the observed data based on date of diagnosis using the delay distribution

```{r}
delay_correct <- ts_date_dx %>%
  mutate(current_date=max(date),
          delay = as.numeric(current_date - date) ) %>%
  left_join(proportion_reporting, by='delay')%>%
  mutate(corrected_cases= N_date_dx/cum_prop)
```

#Plot the reported and corrected cases

```{r}
ggplot(delay_correct, aes(x=date, y=N_date_dx)) +
  geom_line()+
  geom_line( aes(x=date, y=corrected_cases), color='red') +
  theme_classic()

```
compare to actual complete cases (not yet fully observed)

```{r}
 full_ds <- readRDS( 'Data/reported_cases_full_final.rds') %>%
  group_by(date_of_diagnosis) %>%
  summarize(N_full=n())

ggplot(delay_correct, aes(x=date, y=N_date_dx)) +
  geom_line()+
  geom_line( aes(x=date, y=corrected_cases), color='red') +
  geom_line(data=full_ds,aes(x=date_of_diagnosis, y=N_full), color='blue') +
  theme_classic() +
  xlim(as.Date('2019-03-01'),as.Date('2019-03-20'))+
  ylim(0,20)

```







